Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-09-09 article/article.Rmd
Commit 2024-09-09 09:11

```{r loadtrends, echo=F, include=T, warning=FALSE, message=FALSE}
dailytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear.csv"))
dailyCE <- dailytrends[DATA == "ST_E_daily"]
yearlytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear_Proper.csv"))
## regular trends
yearlyCE <- yearlytrends[DATA == "ST_E_yearly"]

@@

@@

The distribution of the number of CE events as a function of OI is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}.  There is an inverse
relationship between the frequency of CE events and OI with an exponential-like
decline. This is expected, as the stronger the CE events are, the rarer the
conditions favoring the occurrence of CE events.

!!!!

For the majority (over $62\,\%$) of
the CE events the OI is below the long-term average of
$`r round(mean(ST_E_yearly$GLB_diff.mean),1)`\,\text{W}/\text{m}^2$, while about
$8.1\,\%$ of the events correspond to OI larger than
$100\,\text{W}/\text{m}^2$ and
up to the highest value of $412.4\,\text{W}/\text{m}^2$.  This distribution is
indicative of the magnitude and the probability of the expected CE events over
Thessaloniki.  Similar distribution of CE events, albeit with larger OI values, has
been reported by @Vamvakas2020, for the city of Patras. This site is located
$2.5^\circ$ south of Thessaloniki and is exposed to air masses coming mainly from the
eastern Mediterranean resulting in different cloud patterns, that may affect the
characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several minutes or even
more than an hour. In order to study the characteristics of these consecutive events,
we grouped them into bins of increasing duration in steps of one minute.  We have
identified `r ST_G0[, .N]` groups of CE in the whole period of study, where the group
of the longest duration of
`r ST_G0[, max(GLB_diff.N)]` minutes occurred on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")` in a SZA range of 
$`r round(ST_G0[which.max(GLB_diff.N), SZA.min],1)`^\circ$ -- 
$`r round(ST_G0[which.max(GLB_diff.N), SZA.max],1)`^\circ$.
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the frequency
distribution of the CE events according to their duration.  We conclude that although
some groups of events last for more than an hour, about $80\,\%$ have duration of
less than 5 minutes.

>>>>

```{r loadtrends, echo=F, include=T, warning=FALSE, message=FALSE}
dailytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear.csv"))
dailyCE <- dailytrends[DATA == "ST_E_daily"]
yearlytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear_Proper.csv"))
## regular trends
# yearlyCE <- yearlytrends[DATA == "ST_E_yearly"]

@@

## some values
some01 <- readRDS("~/MANUSCRIPTS/02_enhancement/data/some_values_01.Rds")

@@

The distribution of the number of CE events as a function of OI is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}.  There is an inverse relationship
between the frequency of CE events and OI with an exponential-like decline. This is
expected, as the stronger the CE events are, the rarer the conditions favouring the
occurrence of CE events.
For the majority (over $`r round(some01$Enh_pc_below_mean,0)`\,\%$) of
the CE events the OI is below the long-term average of
$`r round(some01$Enh_OI_AVG,1)`\,\text{W}/\text{m}^2$, while
$`r round(some01$Enh_pc_below_100,1)`\,\%$ of the events correspond to OI larger than
$100\,\text{W}/\text{m}^2$ and up to the highest value of
$`r round(some01$Max_OI,1)`\,\text{W}/\text{m}^2$.
This distribution is indicative of the magnitude and the probability of the expected
CE events over Thessaloniki.  Similar distribution of CE events, albeit with larger
OI values, has been reported by @Vamvakas2020, for the city of Patras. This site is
located $2.5^\circ$ south of Thessaloniki and is exposed to air masses coming mainly
from the eastern Mediterranean, resulting in different cloud patterns, that may
affect the characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several minutes or even
more than an hour. To study the characteristics of these consecutive events, we
grouped them into bins of increasing duration in steps of one minute.  We have
identified `r ST_G0[, .N]` groups of CE events in the whole period of study, where
the group with the longest duration of
`r ST_G0[, max(GLB_diff.N)]` minutes occurred on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`
with SZA ranging between
$`r round(ST_G0[which.max(GLB_diff.N), SZA.min],1)`^\circ$ and
$`r round(ST_G0[which.max(GLB_diff.N), SZA.max],1)`^\circ$.
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the frequency
distribution of the CE events according to their duration.  We conclude that although
some groups of events last for more than an hour, about $80\,\%$ have a duration of
less than 5 minutes.

($`r ceiling(100*ST_G0[,sum(GLB_ench.N>5)/.N])`\,\%$)

### 2024-09-09 article/article.tex
Commit 2024-09-09 09:11

The distribution of the number of CE events as a function of OI is shown
in Figure\nobreakspace{}\ref{fig:ovir-distribution}. There is an inverse
relationship between the frequency of CE events and OI with an
exponential-like decline. This is expected, as the stronger the CE
events are, the rarer the conditions favoring the occurrence of CE
events.

!!!!

For the majority (over \(62\,\%\)) of the CE events the OI is below the
long-term average of \(42.6\,\text{W}/\text{m}^2\), while about
\(8.1\,\%\) of the events correspond to OI larger than
\(100\,\text{W}/\text{m}^2\) and up to the highest value of
\(412.4\,\text{W}/\text{m}^2\). This distribution is indicative of the
magnitude and the probability of the expected CE events over
Thessaloniki. Similar distribution of CE events, albeit with larger OI
values, has been reported by \citet{Vamvakas2020}, for the city of
Patras. This site is located \(2.5^\circ\) south of Thessaloniki and is
exposed to air masses coming mainly from the eastern Mediterranean
resulting in different cloud patterns, that may affect the
characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several
minutes or even more than an hour. In order to study the characteristics
of these consecutive events, we grouped them into bins of increasing
duration in steps of one minute. We have identified 34849 groups of CE
in the whole period of study, where the group of the longest duration of
140 minutes occurred on 07~July 2013 in a SZA range of \(52.1^\circ\) --
\(77.9^\circ\). Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}
shows the frequency distribution of the CE events according to their
duration. We conclude that although some groups of events last for more
than an hour, about \(80\,\%\) have duration of less than 5 minutes.

>>>>

The distribution of the number of CE events as a function of OI is shown
in Figure\nobreakspace{}\ref{fig:ovir-distribution}. There is an inverse
relationship between the frequency of CE events and OI with an
exponential-like decline. This is expected, as the stronger the CE
events are, the rarer the conditions favouring the occurrence of CE
events. For the majority (over \(62\,\%\)) of the CE events the OI is
below the long-term average of \(42.6\,\text{W}/\text{m}^2\), while
\(9.6\,\%\) of the events correspond to OI larger than
\(100\,\text{W}/\text{m}^2\) and up to the highest value of
\(437\,\text{W}/\text{m}^2\). This distribution is indicative of the
magnitude and the probability of the expected CE events over
Thessaloniki. Similar distribution of CE events, albeit with larger OI
values, has been reported by \citet{Vamvakas2020}, for the city of
Patras. This site is located \(2.5^\circ\) south of Thessaloniki and is
exposed to air masses coming mainly from the eastern Mediterranean,
resulting in different cloud patterns, that may affect the
characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several
minutes or even more than an hour. To study the characteristics of these
consecutive events, we grouped them into bins of increasing duration in
steps of one minute. We have identified 34849 groups of CE events in the
whole period of study, where the group with the longest duration of 140
minutes occurred on 07~July 2013 with SZA ranging between \(52.1^\circ\)
and \(77.9^\circ\).
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the
frequency distribution of the CE events according to their duration. We
conclude that although some groups of events last for more than an hour,
about \(80\,\%\) have a duration of less than 5 minutes.

(\(21\,\%\))
